Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy

### 2021-04-23 manuscript/manuscript.Rmd
Minor edits from @marcdotson

Legal crackdown has also affected how and where funds are allocated. Following the passage of anti-NGO laws, official donors move funds away from contentious causes such as human rights, media, corruption, advocacy, among others, towards tamer and more regime-compatible causes such as health, education [@ChaudhryHeiss:2018]. The latter trend has also been exacerbated by a preference among donors working in the official aid and democracy promotion establishment to gradually increase funding directed towards tame causes, at the expense of more contentious causes. This is because many tame causes can have easily quantifiable, measurable outputs, which makes it easier to show a program's effectiveness over time [@Bush:2015].

We argue that individual private donors can help fill in an important gap. Individual donors do not use the same performance-based metrics as official donors [@desaikharas]. Rather, individual donors often use heuristics to simply their decision-making, making cursory judgments about an organization's issue area, mission, vision, and values and seek out supplementary information from friends, family, and acquaintances [@ChaudhryHeiss:2021; @TBPrakash:2017; @SzperPrakash:2011]. Due to these differences, individual donors can help address some of the challenges arising from reduced public funding available to INGOs. To be clear, we are not arguing that individual giving would be a perfect substitute for NGOs' net loss of income—rather, private donations can be an important supplement when facing reduced funding. Moreover, private philanthropy can also supplement foundation funding, which has been less adversely affected by hostile legal environments [@NeedlesWolcheckSato:2018].

@@

We analyze the effect of different combinations of organizational features and individual respondent attributes with a two-level hierarchical Bayesian multinomial logit model.[^model-details] Simply put, the first level of the model predicts individual donor preferences for various combinations of organizational features (e.g. the combination of human rights issues, financial transparency, government funding, and government crackdown), while the second level predicts the heterogeneity of these fitted preferences as a function of individual donor characteristics (e.g. education, age, and attitudes towards charity, voluntarism, and religion).

@@

To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates simultaneously. Conjoint simulation offers useful analytic benefits. Since we are less concerned with the marginal effects of single donor characteristics and more interested in how constellations of individual and organizational attributes interact with each other, we can vary different combinations of donor and organization characteristics and hold other constant. This allows us not only to see the connections between donor trust and INGO organizational practices and find which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, but also which donor characteristics are associated with the propensity to donate.

@@

The sample of U.S. residents with a prior history of charitable giving also raises questions about generalizability and external validity. Public expectations about the role of government and private funding of the nonprofit sectors differ by country. In the U.S., for instance, NGOs often receive substantial funding through government grants, while those in the UK and France, such as Amnesty International and Médecins Sans Frontières, receive minimal government funding [@Stroup:2012]. This could imply that donors in UK and France may have very different preferences when choosing to donate to government or privately-funded organizations. More comparative research is needed to see if our results would also replicate cross-nationally.

@@

Our results have important implications for NGOs as they navigate a shrinking civic space and consider the use of different frames to tailor their fundraising appeals to individual donors. Our simulation finds that at least for some kinds of personas, NGOs may benefit from publicizing when they are targets of government crackdown, including that information in their fundraising appeals. However, the larger challenge may be addressing the concerns of donors with low levels of trust, as they are the least likely to donate organizations facing criticism and crackdown abroad. As @TBPrakash:2017 argue, "trust in the charity sector cannot be assumed: the sector has to demonstrate that the trust it receives is justified" (p. 643).

>>>>

Legal crackdown has also affected how and where funds are allocated. Following the passage of anti-NGO laws, official donors move funds away from contentious causes such as human rights, media, corruption, advocacy, among others, towards tamer and more regime-compatible causes such as health and education [@ChaudhryHeiss:2018]. The latter trend has also been exacerbated by a preference among donors working in the official aid and democracy promotion establishment to gradually increase funding directed towards tame causes, at the expense of more contentious causes. This is because many tame causes can have easily quantifiable, measurable outputs, which makes it easier to show a program's effectiveness over time [@Bush:2015].

We argue that individual private donors can help fill in an important gap. Individual donors do not use the same performance-based metrics as official donors [@desaikharas]. Rather, individual donors often use heuristics to simplify their decision-making, making cursory judgments about an organization's issue area, mission, vision, and values and seek out supplementary information from friends, family, and acquaintances [@ChaudhryHeiss:2021; @TBPrakash:2017; @SzperPrakash:2011]. Due to these differences, individual donors can help address some of the challenges arising from reduced public funding available to INGOs. To be clear, we are not arguing that individual giving would be a perfect substitute for NGOs' net loss of income—rather, private donations can be an important supplement when facing reduced funding. Moreover, private philanthropy can also supplement foundation funding, which has been less adversely affected by hostile legal environments [@NeedlesWolcheckSato:2018].

@@

We analyze the effect of different combinations of organizational features and individual respondent attributes with a two-level hierarchical Bayesian multinomial logit model.[^model-details] Simply put, the first level of the model predicts individual donor preferences for various combinations of organizational features (e.g. the combination of human rights issues, financial transparency, government funding, and government crackdown) as a function of individual donor characteristics (e.g. education, age, and attitudes towards charity, voluntarism, and religion) while the second level of the model uses these fitted preferences to predict the ultimate choice of donation.

@@

To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates simultaneously. Simulating the philanthropy market offers useful analytic benefits. Since we are less concerned with the marginal effects of single donor characteristics and more interested in how constellations of individual and organizational attributes interact with each other, we can vary different combinations of donor and organization characteristics and hold others constant. This allows us not only to see the connections between donor trust and INGO organizational practices and find which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, but also which donor characteristics are associated with the propensity to donate.

@@

The sample of U.S. residents with a prior history of charitable giving also raises questions about generalizability and external validity. Public expectations about the role of government and private funding of the nonprofit sectors differ by country. In the U.S., for instance, NGOs often receive substantial funding through government grants, while those in the UK and France, such as Amnesty International and Médecins Sans Frontières, receive minimal government funding [@Stroup:2012]. This could imply that donors in the UK and France may have very different preferences when choosing to donate to government or privately-funded organizations. More comparative research is needed to see if our results would also replicate cross-nationally.

@@

Our results have important implications for NGOs as they navigate a shrinking civic space and consider the use of different frames to tailor their fundraising appeals to individual donors. Our simulation finds that at least for some kinds of personas, NGOs may benefit from publicizing when they are targets of government crackdown, including that information in their fundraising appeals. However, the larger challenge may be addressing the concerns of donors with low levels of trust, as they are the least likely to donate to organizations facing criticism and crackdown abroad. As @TBPrakash:2017 argue, "trust in the charity sector cannot be assumed: the sector has to demonstrate that the trust it receives is justified" (p. 643).
